---
title: "Modeling, Plotting and Regression Tables"
author: "Mykola Dereva"
date: "June 11, 2020"
output: github_document
---

```{r include=FALSE}
rm(list = ls())

library(tidyverse)
library(readr)
library(here)
library(janitor)
library(ggeasy)
library(lmtest)
library(sandwich)
library(stargazer)
library(car)
library(robustbase)

# specify tables output. 
# either "text" or "html"
output <- "text"

set.seed(42)
```


load our cleaned data 

```{r}
analysis_full <- readRDS(here("data", "clean data", "Analysis_data.Rds"))
```


## Prepare dependent variable




MPI should be in the range between 0 and 1. in out data it is not the case.
there are few values below zero. 
Let us assume that negative MPI value indicate no market power.
Thus change it to 0. 
And since we cannot take a log of 0. we will add small constant.
Also there is one observation higher than 1, I will drop it. 


```{r}
analysis <- analysis_full %>%
  filter(index < 1,
         # drop observations with conjectural variation
         CV == 0) %>% 
  mutate(index = if_else(index < 0, true = 0, false = index)) %>% 
  mutate(index = index + 0.005)
```


```{r}
summary(analysis$index)
```

Number of articles with MP < 0.05
```{r}
analysis %>% 
  mutate(low_mp = index < 0.5) %>% 
  group_by(low_mp) %>% 
  count()
  
```



## Analysis

Select only columns we need for the analysis.
Rename columns once again. 
And create columns with dummy variables

```{r}
glimpse(analysis)
```




```{r warning=FALSE}
analysis %>%
  summarise_all(~ sum(!is.na(.x))) %>%
  gather(column, not_na)
```





```{r}
log_lm <- lm(log(index) ~ app_PTA + app_GIM 
             #+ obs_n + I(obs_n^2)
             + LI
             + log(obs_n) 
             + freq_Yearly + freq_Monthly      
             + count_n_america + count_europe
             + agg_Firm + agg_Region
             + SC_Processor
             + after_2005 
             + type_Oligopsony 
             + ind_meat + ind_tobacco + ind_beverages + ind_dairy + ind_oils 
             + ind_fruits + ind_cereals + ind_sugar + ind_paper + ind_bread_or_flour
             + perish
             + WOS
             + log(mean_pse)*type_Oligopsony
             #+ sfarm_share_by_cat
             #+ log(mean_pp)
             #+ log(mean_pse)*count_n_america
             #+ edb_index
             #+ sfarm_share
             #+ start_business
                 , 
         data = analysis)

summary(log_lm)
```





```{r}
plot(log_lm)
```

seems that we have hetoroscadasticity issue.
Lets try to use heteroscadasticity robust se and to compare it with the regular

```{r}
log_lm_se <- vcovHC(log_lm, type = "HC")
log_lm_se <- sqrt(diag(log_lm_se))



stargazer(log_lm, log_lm, se = list(NULL, log_lm_se),
          column.labels = c("default SE", "robust SE"), type = "text",
          omit.stat = c("f", "ser", "n"), align = TRUE)
```




```{r}
get_robust_se <- function(lm_model)  {
  matrix <- vcovHC(lm_model, type = "HC0")
  robust_se <- sqrt(diag(matrix))
  return(robust_se)
}

```



## Robust Linear Regression


```{r}
rlm <- lmrob(formula(log_lm),
               data = analysis,
               setting = "KS2014"
               )
summary(rlm)
```



```{r}
plot(rlm)
```







# Summary Tables

I will work with dataset "clean" again

```{r}
clean <- readRDS(here("data", "clean data", "Full_data.Rds"))
```


First of all lets create column with first author and year
to uniquely identify an article
```{r}
clean <- clean %>%
  mutate(article_ref = str_c(
           str_match(authors, pattern = "^\\w[’]*\\w+" ),
           " (", year, ")")
         ) 

head(clean$article_ref, 20)
```


summary table of all articles used in analysis




### Table 1. List of studies
```{r}
table_1 <- clean %>%
  group_by(article_ref, approach, SC_level, mp_type, data_freq, period) %>%
  summarise(N = n()) %>%
  adorn_totals("row")

head(table_1)
```

```{r}
table_1 %>%
  stargazer(type = output,
            style = "aer",
            title = "List of studies used",
            summary = FALSE, rownames = FALSE,
            out = "Tables/1.List of studies.doc"
            )
```


### Table 2. Number of observations by country

```{r}
table_2 <-  clean %>%
  mutate(country = as_factor(country) %>%
           fct_infreq() %>%
           fct_lump_min(min = 10) 
         ) %>%
  group_by(country) %>%
  summarise(n = n(), 
            mean_mp = round(mean(mp_index), 4)
            ) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n) * 100, 2) ) %>%
  select(country, n, prop, mean_mp) %>%
  adorn_totals("row")
```

```{r}
table_2 %>%
  stargazer(type = output,
            style = "aer",
            title = "Summary of MP observations by country",
            summary = FALSE, rownames = FALSE,
            out = "Tables/2.List of countries.doc")
```

Visualize the distribution

```{r}
clean %>%
    mutate(country = as_factor(country) %>%
           fct_infreq() %>%
           fct_lump_min(min = 10) 
         ) %>%
  ggplot(aes(x = country,
             y = mp_index,
             color = approach)) + 
  geom_jitter(width = 0.25, height = 0, alpha = 0.6, size = 2) +
  theme_minimal() +
  easy_legend_at("bottom") +
  easy_remove_x_axis("title") +
  labs(y = "Market Power Index",
       title = "The distribution of MP indices by country",
       caption = "Source: own representation",
       color = "Approach:") 
```

```{r include=FALSE}
ggsave(filename = "Plots/2. Country distribution plot.png",
       dpi = "print",
       height = 10, width = 15, units = "cm")
```



### Table 3 Number of observations by sector


```{r}
table_3 <- clean %>%
  mutate(industry_fct = fct_infreq(industry_fct) %>%
           fct_relevel("other", after = Inf)
         ) %>%
  group_by(industry_fct) %>%
  summarise(n = n(), 
            mean_mp = round(mean(mp_index), 4)
            ) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n) * 100, 2),
         industry_fct = str_to_title(industry_fct)) %>%
  select(industry_fct, n, prop, mean_mp) %>%
  adorn_totals("row")

```


Save table
```{r}
table_3 %>%
  stargazer(type = output,
            style = "aer",
            title = "Summary of MP observations by sector",
            summary = FALSE, rownames = FALSE,
            out = "Tables/3. List of industries.doc")
```


Visualize the distribution

```{r}
clean %>%
  ggplot(aes(x = fct_reorder(industry_fct, mp_index, median) %>%
               str_to_sentence(),
             y = mp_index,
             color = approach)) + 
  geom_jitter(width = 0.25, height = 0, alpha = 0.6, size = 2) +
  theme_minimal() +
  easy_legend_at("bottom") +
  easy_remove_x_axis("title") +
  labs(y = "Market Power Index",
       title = "The distribution of MP indices by sector",
       caption = "Source: own representation",
       color = "Approach:") 
```
Save plot 
```{r include=FALSE}
ggsave(filename = "Plots/3. Sector distribution plot.png",
       dpi = "print",
       height = 10, width = 15, units = "cm")
```





### Table 4: Number of observations by methodology

```{r}
table_4 <- clean %>%
  group_by(approach) %>%
  summarise(n = n(), 
            mean_mp = round(mean(mp_index), 4)
            ) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n) * 100, 2)) %>%
  select(approach, n, prop, mean_mp) %>%
  adorn_totals("row")
```


```{r}
table_4 %>%
  stargazer(type = output,
            style = "aer",
            title = "Summary of MP observations by approach used",
            summary = FALSE, rownames = FALSE,
            out = "Tables/4. List of Methods.doc")
```



### Table 5: List of observarion frequency

```{r}
table_5 <- clean %>%
  group_by(data_freq) %>%
  summarise(n = n(), 
            mean_mp = round(mean(mp_index), 4)
            ) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n) * 100, 2)) %>%
  select(data_freq, n, prop, mean_mp) %>%
  arrange(-n) %>%
  adorn_totals("row")
```

```{r}
table_5 %>%
  stargazer(type = output,
            style = "aer",
            title = "Summary of MP observations by frequency",
            summary = FALSE, rownames = FALSE,
            out = "Tables/5. List of observarion frequency.doc")
```



### Table 6: Discriptive statistics of variables 

```{r }
analysis %>%
  as.data.frame() %>%
  stargazer(type = output,
            style = "aer",
            align = TRUE, nobs = FALSE,
            out = "Tables/6. Descriprive stat.doc")
```



### Table 7: Regression table




```{r}
model_1 <- lm(log(index) ~ app_PTA + app_GIM +
               after_2005 + type_Oligopsony, 
         data = analysis)
se_1 <- get_robust_se(model_1)

model_2 <- lm(log(index) ~ app_PTA + app_GIM +
               after_2005 + type_Oligopsony +
               log(obs_n) +
               freq_Yearly +
               count_n_america + count_europe, 
         data = analysis)
se_2 <- get_robust_se(model_2)


model_3 <- lm(log(index) ~ app_PTA + app_GIM +
               after_2005 + type_Oligopsony +
               log(obs_n) +
               freq_Yearly +
               count_n_america + count_europe +
               ind_meat + ind_tobacco + ind_beverages + ind_dairy + ind_oils +
               ind_fruits + ind_cereals, 
         data = analysis)
se_3 <- get_robust_se(model_3)

model_4 <- lm(log(index) ~ app_PTA + app_GIM +
                after_2005 + type_Oligopsony +
                log(obs_n) +
                freq_Yearly +
                count_n_america + count_europe +
                ind_meat + ind_tobacco + ind_beverages + ind_dairy + ind_oils +
                ind_fruits + ind_cereals + perish +
                log(mean_pp) + log(mean_pse),
         data = analysis)
se_4 <- get_robust_se(model_4)

model_5 <- lm(log(index) ~ app_PTA + app_GIM +
                after_2005 + type_Oligopsony +
                log(obs_n) +
                freq_Yearly +
                count_n_america + 
                ind_meat + ind_tobacco + ind_beverages + ind_dairy + ind_oils +
                ind_fruits + ind_cereals + perish +
                log(mean_pp) + log(mean_pse) +
                log(sfarm_share) + log(start_business),
         data = analysis)
se_5 <- get_robust_se(model_5)

ind.var.names <- c("Intercept", "PTA", "GIM", "After 2005", "Oligopsony",
                   "Obs. Num. (log)", "Yearly", "USA", "Europe",
                   "Meat", "Tobacco", "Beverages", "Dairy", "Oils", "Fruits",
                   "Cereals", "Perishable", "PP (log)",
                   "PSE (log)", "Small Farm Share (log)",
                   "Start Business Ease (log)")

stargazer(model_1, model_2, model_3, model_4, model_5,
          type = output,
          style = "io",
          omit.stat = c("rsq"), 
          title = "Regression Results",
          dep.var.labels = "Market Power Index (log)",
          covariate.labels = ind.var.names,
          se = list(se_1, se_2, se_3, se_4, se_5),      #  Robust se
          align = TRUE,
          df = FALSE,
          intercept.bottom = FALSE,
          out = "Tables/7. Regression Table.doc"
          )
```



### Table 8: Methods comparison 




```{r}
clean %>%
  ggplot(aes(x = n_of_obs, y = mp_index, color = approach)) +
  geom_point(size = 4, alpha = 0.8, shape = 1) +
  scale_x_continuous(trans = "log10") +
  theme_minimal() +
  easy_move_legend(to = "bottom") +
    labs(x = "Number of observations (log10)",
         y = "Market Power Index",
         title = "Relationship between MP index and number of observations \nused in a study",
         caption = "Source: own representation") +
  easy_add_legend_title("Approach used:")
```

```{r include=FALSE}
ggsave(filename = "Plots/4. Observarion number plot.png",
       dpi = "print",
       height = 10, width = 15, units = "cm")
```





### Tables PP and PSE

```{r include=FALSE}
model_1 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pse), 
         data = analysis)
se_1 <- get_robust_se(model_1)

model_2 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pp), 
         data = analysis)
se_2 <- get_robust_se(model_2)

model_3 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pp) + log(mean_pse), 
         data = analysis)
se_3 <- get_robust_se(model_3)


model_4 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                mean_pse, 
         data = analysis)
se_4 <- get_robust_se(model_4)

model_5 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                mean_pp, 
         data = analysis)
se_5 <- get_robust_se(model_5)

model_6 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                mean_pp + mean_pse, 
         data = analysis)
se_6 <- get_robust_se(model_6)
```

```{r }
stargazer(model_1, model_2, model_3, model_4, model_5, model_6,
          type = output,
          style = "aer",
          #title = "Regressions Results",
          keep.stat = c("n", "rsq", "adj.rsq"), 
          title = "Analysis Results",
          #column.labels =  c("Model 1", "Model 2", "Model 3"),
          dep.var.labels = "Market Power Index (log transformed)",
          se = list(se_1, se_2, se_3, se_4, se_5, se_6),
          align = TRUE,
          intercept.bottom = FALSE,
          out = "Tables/9. PP vs PSE.doc"
          )
```

### tables small farm and perish

```{r include=FALSE}
model_1 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pse) + perish, 
         data = analysis)
se_1 <- get_robust_se(model_1)

model_2 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                #ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pse) + perish, 
         data = analysis)
se_2 <- get_robust_se(model_2)

model_3 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(sfarm_share), 
         data = analysis)
se_3 <- get_robust_se(model_3)

model_4 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(sfarm_share_by_cat), 
         data = analysis)
se_4 <- get_robust_se(model_4)
```

```{r}
stargazer(model_1, model_2, model_3, model_4,
          type = output,
          style = "aer",
          #title = "Regressions Results",
          keep.stat = c("n", "rsq", "adj.rsq"), 
          title = "Analysis Results",
          #column.labels =  c("Model 1", "Model 2", "Model 3"),
          dep.var.labels = "Market Power Index (log transformed)",
          se = list(se_1, se_2, se_3),
          align = TRUE,
          intercept.bottom = FALSE,
          out = "Tables/10. Small farm and perish.doc"
          )
```



### tables doing business

```{r include=FALSE}
model_1 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pse), 
         data = analysis)
se_1 <- get_robust_se(model_1)

model_2 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pse) + start_business, 
         data = analysis)
se_2 <- get_robust_se(model_2)

model_3 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                #log(mean_pse) +
                log(start_business), 
         data = analysis)
se_3 <- get_robust_se(model_3)
```


```{r}
stargazer(model_1, model_2, model_3,
          type = output,
          style = "aer",
          #title = "Regressions Results",
          keep.stat = c("n", "rsq", "adj.rsq"), 
          title = "Analysis Results",
          #column.labels =  c("Model 1", "Model 2", "Model 3"),
          dep.var.labels = "Market Power Index (log transformed)",
          se = list(se_1, se_2, se_3),
          align = TRUE,
          intercept.bottom = FALSE,
          out = "Tables/11. DB.doc"
          )
```

## Table conference Rajske Teplice
```{r}
model_1 <- lm(log(index) ~ app_PTA + app_GIM +
               after_2005 + type_Oligopsony, 
         data = analysis)
se_1 <- get_robust_se(model_1)

model_2 <- lm(update(formula(model_1), ~ . +
                       log(obs_n) + freq_Yearly +
                       count_n_america + count_europe),

         data = analysis)
se_2 <- get_robust_se(model_2)


model_3 <- lm(update(formula(model_2), ~ . +
               ind_meat + ind_tobacco + ind_beverages + ind_dairy + ind_oils +
               ind_fruits + ind_cereals), 
         data = analysis)
se_3 <- get_robust_se(model_3)


model_4 <- lm(update(formula(model_3), ~ . +
                       perish),
                       data = analysis)
se_4 <- get_robust_se(model_4)

model_5 <- MASS::rlm(formula(model_4),
                     data = analysis_full,
                     maxit = 100,
                     psi = "psi.bisquare")
se_5 <- get_robust_se(model_5)

ind.var.names <- c("Intercept", "PTA", "GIM", "After 2005", "Oligopsony",
                   "Obs. Num. (log)", "Yearly", "USA", "Europe",
                   "Meat", "Tobacco", "Beverages", "Dairy", "Oils", "Fruits",
                   "Cereals", "Perishable")

stargazer(model_1, model_2, model_3, model_4, model_5,
          type = output,
          style = "io",
          title = "Regression Results",
          dep.var.labels = "Market Power Index (log)",
          covariate.labels = ind.var.names,
          dep.var.labels.include = FALSE,
          omit.stat = c("adj.rsq"),
          se = list(se_1, se_2, se_3, se_4, se_5),      #  Robust se
          align = TRUE,
          df = FALSE,
          intercept.bottom = FALSE,
          star.cutoffs = c(0.05, 0.01, 0.001),
          out = "Tables/Regression Table Teplice.doc"
          )
```

