---
title: "Modeling, Plotting and Regression Tables"
author: "Mykola Dereva"
date: "June 11, 2020"
output: github_document
---

```{r include=FALSE}
rm(list=ls())

library(tidyverse)
library(readr)
library(here)
library(janitor)
library(ggeasy)
library(lmtest)
library(sandwich)
library(stargazer)
library(car)

set.seed(42)
```


load our cleaned data 

```{r}
clean <- readRDS(here("data", "clean data", "Clean_data.Rds"))
```


## Analysis

Select only columns we need for the analysis.
Rename columns once again. 
And create columns with dummy variables

```{r}
glimpse(clean)
```


```{r}
library(fastDummies)

analysis <- clean %>%
  select(approach, data_freq, mp_type, mp_index, sfarm_share,
         industry_fct, country_fct, n_of_obs, after_2005, perish,
         mean_pse, mean_pp, start_business) %>%
  rename(app = approach,
         freq = data_freq,
         type = mp_type,
         index = mp_index,
         ind = industry_fct,
         count = country_fct,
         obs_n = n_of_obs) %>%
  dummy_columns(select_columns = c("freq", "type", "ind", "count", "app"),
                remove_selected_columns = TRUE)
```

final quick check of the data
```{r}
summary(analysis)
```


```{r}
analysis %>%
  summarise_all(funs(sum(is.na(.)))) %>%
  gather(column, n_na)
```






```{r}
summary(lm(index ~ ., data=analysis))
```

lets try automatic variable selection
```{r}
library(leaps)
```

```{r}
auto.select <- regsubsets(log(index) ~ ., data=analysis,
                          method = "exhaustive", nvmax = 20)
```

```{r}
x <- summary(auto.select)

plot(x$bic, main="BIC")
plot(x$adjr2, main="Adj R^2")
```


So according BIC the oprimal number of variables is around 3
While according to adj. R^2 - around 8-10

```{r}
plot(auto.select)
```


```{r}
log_lm <- lm(log(index) ~ app_PTA + app_GIM +
                 obs_n + 
                 count_n_america +
                 after_2005 +
                 type_Oligopsony +
                 ind_dairy + ind_meat + ind_cereals + ind_beverages + ind_oils +
                 mean_pp + mean_pse, 
         data=analysis)
summary(log_lm)
```

```{r}
bptest(log_lm)
```


```{r}
plot(log_lm)
```

seems that we have hetoroscadasticity issue.
Lets try to use heteroscadasticity robust se and to compare it with the regular

```{r}
log_lm_se <- vcovHC(log_lm, type = "HC")
log_lm_se <- sqrt(diag(log_lm_se))
log_lm_se
```


```{r}
stargazer(log_lm, log_lm, se=list(NULL, log_lm_se),
          column.labels=c("default SE","robust SE"), type = "text",
          omit.stat=c("f", "ser", "n"), align = TRUE)
```



#### Without Log transformation
```{r}
lm <- lm(index ~ app_PTA + app_GIM +
                 obs_n + 
                 count_n_america +
                 after_2005 +
                 type_Oligopsony +
                 ind_dairy + ind_meat + ind_cereals + ind_beverages + ind_oils, 
         data=analysis)
summary(lm)
```

```{r}
bptest(lm)
```

```{r}
plot(lm)
```

```{r}
get_robust_se <- function(lm_model)  {
  matrix <- vcovHC(lm_model, type = "HC0")
  robust_se <- sqrt(diag(matrix))
  return(robust_se)
}

get_robust_se(lm)
```



```{r}
model_1 <- lm(log(index) ~ app_PTA + app_GIM + type_Oligopoly, 
                        data=analysis)
se_1 <- get_robust_se(model_1)

model_2 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + after_2005 + 
                freq_Yearly + freq_Quaterly +
                count_n_america + count_europe, 
         data=analysis)
se_2 <- get_robust_se(model_2)

model_3 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + after_2005 + 
                freq_Yearly + freq_Quaterly +
                count_n_america + count_europe +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils, 
         data=analysis)
se_3 <- get_robust_se(model_3)


model_4 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pp) + log(mean_pse), 
         data=analysis)
se_4 <- get_robust_se(model_4)
```

#### Test for heteroscadasticity
```{r}
bptest(model_1)
```
```{r}
bptest(model_2)
```

```{r}
bptest(model_3)
```

```{r}
bptest(model_4)
```

We can see that all 3 models expesed to the heteroscadasticity issue. 


## Summary tables
```{r}
stargazer(model_1, model_2, model_3, model_4,
          type = "text",
          keep.stat=c("n", "rsq", "adj.rsq"), 
          title = "Analysis Results",
          #column.labels =  c("Model 1", "Model 2", "Model 3"),
          dep.var.labels = "Market Power Index (log transformed)",
          se = list(se_1, se_2, se_3, se_4),
          align = TRUE,
          intercept.bottom = FALSE
          )
```

### F-test 

In the regressions results neither of dummmies regarding frequency of observation was significant. 
To be completely sure that they are do not have any influence on the MPI lets make an F test. 

```{r}
F.test.freq <-  linearHypothesis(model_3, c("freq_Yearly", "freq_Quaterly"), 
                                 white.adjust = "hc1") # Heteroscad. robust F test
F.test.freq
```
According to F test we cannot reject H:0 

Lets check if type there influence of industry on the MPI 
```{r}
industry <- matchCoefs(model_3, "ind") # extract all industry dummies
F.test.ind <-  linearHypothesis(model_3, industry, white.adjust = "hc1")
F.test.ind
```
Therefore we reject 0 hypothesis that all types of industry have zero coeficient. 




### Determinants of each approach 


Finally lets make a linear model for each approahc and check if there is any significant differences in coefitients between them

```{r}
GIM <- lm(log(index) ~ type_Oligopoly +
                obs_n + after_2005 + 
                freq_Yearly + freq_Quaterly +
                count_n_america + count_europe +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils, 
         data= subset(analysis, app_GIM == 1)) 

PTA <- lm(log(index) ~ type_Oligopoly +
                obs_n + after_2005 + 
                freq_Yearly + freq_Quaterly +
                count_n_america + count_europe +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils, 
         data= subset(analysis, app_PTA == 1))

SFA <- lm(log(index) ~ type_Oligopoly +
                obs_n + after_2005 + 
                freq_Yearly + freq_Quaterly +
                count_n_america + count_europe +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils, 
         data= subset(analysis, app_SFA == 1)) 
```


```{r}
bptest(SFA)
```
```{r}
bptest(PTA)
```

```{r}
bptest(GIM)
```

According to bptest there is no evidence of heteroscadasticity, so we can use default SE


```{r}
stargazer(GIM, PTA, SFA,
          type = "text",
          omit.stat=c("f", "ser"), 
          title = "Analysis Results",
          column.labels =  c("GIM", "PTA", "SFA"),
          dep.var.labels = "Market Power Index (log transformed)",
          align = TRUE,
          intercept.bottom = FALSE
          )
```


I think I am done with the analysis


# Summary Tables

I will work with dataset "clean" again

First of all lets create column with fitst author and year
to uniquely identify an article
```{r}
clean <- clean %>%
  mutate(article_ref = str_c(
           str_match(authors, pattern = "^\\w[’]*\\w+" ),
           " (", year, ")")
         ) 

head(clean$article_ref, 20)
```


summary table of all articles used in analysis



```{r}
#replace misspeled value
clean <- clean %>%
  mutate(data_freq = str_replace(data_freq, "Quaterly", "Quarterly")) 
```

### Table 1. List of studies
```{r}
table_1 <- clean %>%
  group_by(article_ref, approach, mp_type, data_freq, period) %>%
  summarise(N = n()) %>%
  adorn_totals("row")

head(table_1)
```

```{r include=FALSE}
table_1 %>%
  stargazer(type="html",
            style = "aer",
            title = "List of studies used",
            summary = FALSE, rownames = FALSE,
            out = "Tables/1.List of studies.doc"
            )
```


### Table 2. Number of observations by country

```{r}
table_2 <-  clean %>%
  mutate(country = as_factor(country) %>%
           fct_infreq() %>%
           fct_lump_min(min = 3) 
         ) %>%
  group_by(country) %>%
  summarise(n = n(), 
            mean_mp = round(mean(mp_index), 4)
            ) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n) * 100, 2) ) %>%
  select(country, n, prop, mean_mp) %>%
  adorn_totals("row")

table_2
```

```{r include=FALSE}
table_2 %>%
  stargazer(type="html",
            style = "aer",
            title = "Summary of MP observations by country",
            summary = FALSE, rownames = FALSE,
            out = "Tables/2.List of countries.doc")
```

Visualize the distribution

```{r}
clean %>%
    mutate(country = as_factor(country) %>%
           fct_infreq() %>%
           fct_lump_min(min = 4) 
         ) %>%
  ggplot(aes(x = country,
             y = mp_index,
             color = approach)) + 
  geom_jitter(width = 0.25, height = 0, alpha = 0.6, size = 2) +
  theme_minimal() +
  easy_legend_at("bottom") +
  easy_remove_x_axis("title") +
  labs(y = "Market Power Index",
       title = "The distribution of MP indices by country",
       caption = "Source: own representation",
       color = "Approach:") 
```
```{r}
ggsave(filename = "Plots/2. Country distribution plot.png",
       dpi = "print",
       height = 10, width = 15, units = "cm")
```



### Table 3 Number of observations by sector


```{r}
table_3 <- clean %>%
  mutate(industry_fct = fct_infreq(industry_fct) %>%
           fct_relevel("other", after = Inf)
         ) %>%
  group_by(industry_fct) %>%
  summarise(n = n(), 
            mean_mp = round(mean(mp_index), 4)
            ) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n) * 100, 2),
         industry_fct = str_to_title(industry_fct)) %>%
  select(industry_fct, n, prop, mean_mp) %>%
  adorn_totals("row")

table_3
```


Save table
```{r include=FALSE}
table_3 %>%
  stargazer(type="html",
            style = "aer",
            title = "Summary of MP observations by sector",
            summary = FALSE, rownames = FALSE,
            out = "Tables/3. List of industries.doc")
```


Visualise the distribution
```{r}
clean %>%
  ggplot(aes(x = fct_reorder(industry_fct, mp_index, median) %>%
               str_to_sentence(),
             y = mp_index,
             color = approach)) + 
  geom_jitter(width = 0.25, height = 0, alpha = 0.6, size = 2) +
  theme_minimal() +
  easy_legend_at("bottom") +
  easy_remove_x_axis("title") +
  labs(y = "Market Power Index",
       title = "The distribution of MP indices by sector",
       caption = "Source: own representation",
       color = "Approach:") 
```
Save plot 
```{r}
ggsave(filename = "Plots/3. Sector distribution plot.png",
       dpi = "print",
       height = 10, width = 15, units = "cm")
```





### Table 4: Number of observations by methodology

```{r}
table_4 <- clean %>%
  group_by(approach) %>%
  summarise(n = n(), 
            mean_mp = round(mean(mp_index), 4)
            ) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n) * 100, 2)) %>%
  select(approach, n, prop, mean_mp) %>%
  adorn_totals("row")

table_4
```


```{r include=FALSE}
table_4 %>%
  stargazer(type="html",
            style = "aer",
            title = "Summary of MP observations by approach used",
            summary = FALSE, rownames = FALSE,
            out = "Tables/4. List of Methods.doc")
```



### Table 5: List of observarion frequency

```{r}
table_5 <- clean %>%
  group_by(data_freq) %>%
  summarise(n = n(), 
            mean_mp = round(mean(mp_index), 4)
            ) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n) * 100, 2)) %>%
  select(data_freq, n, prop, mean_mp) %>%
  arrange(-n) %>%
  adorn_totals("row")

table_5
```

```{r include=FALSE}
table_5 %>%
  stargazer(type="html",
            style = "aer",
            title = "Summary of MP observations by frequency",
            summary = FALSE, rownames = FALSE,
            out = "Tables/5. List of observarion frequency.doc")
```



### Table 6: Discriptive statistics of variables 

```{r include=FALSE}
analysis %>%
  as.data.frame() %>%
  stargazer(type="html",
            style = "aer",
            align = TRUE, nobs = FALSE,
            out = "Tables/6. Descriprive stat.doc")
```



### Table 7: Regresiion table

```{r}
ind.var.names <- c("Intercept", "PTA", "GIM", "Oligopoly",
                   "Observations Num.", "After 2005", "Yearly",
                   "Quarterly", "USA", "Europe", "Dairy", "Meat",
                   "Beverages", "Cereals", "Oils")
```


```{r include=FALSE}
stargazer(model_1, model_2, model_3,
          type = "html",
          style = "aer",
          #omit.stat=c("f", "ser", "n"), 
          title = "Regression Results",
          dep.var.labels = "Market Power Index (log transformed)",
          covariate.labels = ind.var.names,
          se = list(se_1, se_2, se_3),      #  Robust se
          align = TRUE,
          df = FALSE,
          intercept.bottom = FALSE,
          out = "Tables/7. Regression Table.doc"
          )
```

### Table 8: Methods comparison 

```{r include=FALSE}
stargazer(PTA, GIM, SFA,
          type = "html",
          style = "aer",
          #omit.stat=c("f", "ser", "n"), 
          title = "Methods Comparison",
          dep.var.labels = "Market Power Index",
          column.labels =  c("PTA", "GIM", "SFA"),
          covariate.labels = ind.var.names,
          align = TRUE,
          df = FALSE,
          intercept.bottom = FALSE,
          out = "Tables/8. Method Comparison.doc"
          )
```

### Table 9: F tests industry

```{r include=FALSE}
stargazer(F.test.ind,
          type = "html",
          style = "aer",
          out = "Tables/9. F test industry.doc")
```


```{r}
clean %>%
  ggplot(aes(x=n_of_obs, y=mp_index, color = approach)) +
  geom_point(size = 4, alpha = 0.8, shape = 1) +
  scale_x_continuous(trans = "log10") +
  theme_minimal() +
  easy_move_legend(to = "bottom") +
    labs(x = "Number of observations (log10)",
         y = "Market Power Index",
         title = "Relationship between MP index and number of observations \nused in a study",
         caption = "Source: own representation") +
  easy_add_legend_title("Approach used:")
```

```{r}
ggsave(filename = "Plots/4. Observarion number plot.png",
       dpi = "print",
       height = 10, width = 15, units = "cm")
```





### Tables PP and PSE

```{r}
model_1 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pse), 
         data=analysis)
se_1 <- get_robust_se(model_1)

model_2 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pp), 
         data=analysis)
se_2 <- get_robust_se(model_2)

model_3 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pp) + log(mean_pse), 
         data=analysis)
se_3 <- get_robust_se(model_3)


model_4 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                mean_pse, 
         data=analysis)
se_4 <- get_robust_se(model_4)

model_5 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                mean_pp, 
         data=analysis)
se_5 <- get_robust_se(model_5)

model_6 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                mean_pp + mean_pse, 
         data=analysis)
se_6 <- get_robust_se(model_6)
```
```{r message=FALSE, include=FALSE}
stargazer(model_1, model_2, model_3, model_4, model_5, model_6,
          type = "html",
          style = "aer",
          #title = "Regressions Results",
          keep.stat=c("n", "rsq", "adj.rsq"), 
          title = "Analysis Results",
          #column.labels =  c("Model 1", "Model 2", "Model 3"),
          dep.var.labels = "Market Power Index (log transformed)",
          se = list(se_1, se_2, se_3, se_4, se_5, se_6),
          align = TRUE,
          intercept.bottom = FALSE,
          out = "Tables/9. PP vs PSE.doc"
          )
```

### tables small farm and perish

```{r}
model_1 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pse), 
         data=analysis)
se_1 <- get_robust_se(model_1)

model_2 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                #ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pse) + perish, 
         data=analysis)
se_2 <- get_robust_se(model_2)

model_3 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(sfarm_share) + log(mean_pse), 
         data=analysis)
se_3 <- get_robust_se(model_3)
```

```{r message=FALSE, include=FALSE}
stargazer(model_1, model_2, model_3,
          type = "html",
          style = "aer",
          #title = "Regressions Results",
          keep.stat=c("n", "rsq", "adj.rsq"), 
          title = "Analysis Results",
          #column.labels =  c("Model 1", "Model 2", "Model 3"),
          dep.var.labels = "Market Power Index (log transformed)",
          se = list(se_1, se_2, se_3),
          align = TRUE,
          intercept.bottom = FALSE,
          out = "Tables/10. Small farm and perish.doc"
          )
```



### tables doing business

```{r}
model_1 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pse), 
         data=analysis)
se_1 <- get_robust_se(model_1)

model_2 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                log(mean_pse) + start_business, 
         data=analysis)
se_2 <- get_robust_se(model_2)

model_3 <- lm(log(index) ~ app_PTA + app_GIM +
                type_Oligopoly +
                obs_n + 
                count_n_america +
                ind_dairy + ind_meat + ind_beverages + ind_cereals + ind_oils +
                #log(mean_pse) +
                log(start_business), 
         data=analysis)
se_3 <- get_robust_se(model_3)
```


```{r message=FALSE, include=FALSE}
stargazer(model_1, model_2, model_3,
          type = "html",
          style = "aer",
          #title = "Regressions Results",
          keep.stat=c("n", "rsq", "adj.rsq"), 
          title = "Analysis Results",
          #column.labels =  c("Model 1", "Model 2", "Model 3"),
          dep.var.labels = "Market Power Index (log transformed)",
          se = list(se_1, se_2, se_3),
          align = TRUE,
          intercept.bottom = FALSE,
          out = "Tables/11. DB.doc"
          )
```


